Splitting it up: the \pkg{spduration} split-population duration regression package

### 2015-12-18 manuscript.Rmd
added King and Zeng
added King and Zeng cites
catching up

---
author:
  - name: Andreas Beger
    affiliation: Duke University
    address: >
      Department of Political Science, Duke University
      Durham, NC, USA
    email: ab428@duke.edu
    url: something
  - name: Daniel W. Hill, Jr.
    affiliation: University of Georgia
    address: >
      First line
      Athens, GA, USA
    email: dwhill@uga.edu
  - name: Nils W. Metternich
    affiliation: University College London
    address: >
      First line
      London, UK
    email: n.metternich@ucl.ac.uk
  - name: Michael D. Ward
    affiliation: Duke University
    address: >
      Department of Political Science, Duke University
      Durham, NC, USA
    email: mw160@duke.edu
title:
  formatted: "Splitting it up: the \\pkg{spduration} split-population duration regression package"
  # If you use tex in the formatted title, also supply version without
  plain:     "Splitting it up: the spduration split-population duration regression package"
  # For running headers, if needed
  short:     "\\pkg{spduration}: Split-population duration regression"
abstract: >
  We present an implementation of split-population duration regression in the   \pkg{spduration} R package and an application to data on military coups. The statistical model accounts for units that are immune to a certain outcome and not part of the duration process the researcher is primarily interested in. We provide insights that if immune units exist, we can significantly increase the predictive performance compared to standard duration models. The package includes estimation and several post-estimation methods for split-population Weibull and Loglogistic models.
keywords:
  # at least one keyword must be supplied
  formatted: [regression, rare events, "\\proglang{R}"]
  plain:     [regression, rare events, R]
preamble: >
  \usepackage{amsmath}
  \usepackage{booktabs}
output: 
  rticles::jss_article
bibliography: spdur.bib
---

@@

Many outcomes of interest are rare events. For example, coups, war onset, or mass killings are exceedingly rare events when considering that most countries do not experience high levels of violence. Figure \ref{rates} shows a few examples of published research that models binary outcomes. @fearon2003ethnicity is a widely cited study of civil war onset that uses yearly observations of all major countries; @beger2014ensemble is an example of the positive rates when moving to monthly data for a similar set of countries. The positive rates range from 5 to less than 0.2\% of all data points.

>>>>

---
author:
  - name: Andreas Beger
    affiliation: Duke University
    address: >
      Department of Political Science, Duke University
      Durham, NC, USA
    email: ab428@duke.edu
    url: something
  - name: Daniel W. Hill, Jr.
    affiliation: University of Georgia
    address: >
      Department of International Affairs, University of Georgia
      Athens, GA, USA
    email: dwhill@uga.edu
  - name: Nils W. Metternich
    affiliation: University College London
    address: >
      Department of Political Science, University College London
      London, UK
    email: n.metternich@ucl.ac.uk
  - name: Shahryar Minhas
    affiliation: Duke University
    address: >
      Department of Political Science, Duke University
      Durham, NC
    email: sfm12@duke.edu
  - name: Michael D. Ward
    affiliation: Duke University
    address: >
      Department of Political Science, Duke University
      Durham, NC, USA
    email: mw160@duke.edu
title:
  formatted: "Splitting it up: the \\pkg{spduration} split-population duration regression package"
  # If you use tex in the formatted title, also supply version without
  plain:     "Splitting it up: the spduration split-population duration regression package"
  # For running headers, if needed
  short:     "\\pkg{spduration}: Split-population duration regression"
abstract: >
  We present an implementation of split-population duration regression in the   \pkg{spduration} R package and an application to data on military coups. The statistical model accounts for units that are immune to a certain outcome and not part of the duration process the researcher is primarily interested in. We provide insights that if immune units exist, we can significantly increase the predictive performance compared to standard duration models. The package includes estimation and several post-estimation methods for split-population Weibull and Loglogistic models.
keywords:
  # at least one keyword must be supplied
  formatted: [regression, rare events, "\\proglang{R}"]
  plain:     [regression, rare events, R]
preamble: >
  \usepackage{amsmath}
  \usepackage{booktabs}
output: 
  rticles::jss_article
bibliography: spdur.bib
---

@@

Many outcomes of interest are rare events. For example, coups, war onset, or mass killings are exceedingly rare events when considering that most countries do not experience high levels of violence. Figure \ref{rates} shows a few examples of published research that models binary outcomes. @fearon2003ethnicity is a widely cited study of civil war onset that uses yearly observations of all major countries; @beger2014ensemble is an example of the positive rates when moving to monthly data for a similar set of countries. The positive rates range from 5 to less than 0.2\% of all data points.\footnote{For a discussion of the difficulties rare events can pose for prediction see \citet{king2001logistic} and \citet{king2001explaining}.}
